Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-07-19 GNproofcolumns.tex
Added a new document with a new version of the proof

\usepackage{amssymb, amsmath, verbatim, amsthm,url, multirow,fullpage,mathtools}
\usepackage{longtable, rotating,makecell,array}
\usepackage[aligntableaux=top]{ytableau}

@@

We first note that if $a_{m-1}$ and $a_{m}$ are not consecutive vertices for some $m$, then for any intermediate vertex $a_{m-1}<_i x <_i a_{m}$ we must have $\Prop(x) \subseteq \cP_{m-1}$. This is because if any propagator supported on $x$ is still unassigned when the algorithm reaches $x$, we would have $x \in I_i$. We also note that when $p_m$ contributes $I_i(p_m) = a_m$ to the Grassmann necklace, it must be the clockwise-most propagator (amongst those not yet assigned) that is supported on $a_m$ i.e. clockwise-most amongst propagators in the set $\cP \setminus \cP_{m-1}$.

Now, let $J = \{j_1 <_i j_2<_i \dots <_i j_k\}$ be a $k$-set, and suppose that $J<_i I_i$ in the $<_i$-lexicographic order. Let $m$ be the minimal index such that $j_m <_i a_m$, i.e. the first entry at which $J$ and $I_i$ differ. We will show that $J$ must be a dependent set.

If $\Prop(j_m) = \emptyset$ then $J$ is automatically dependent since the corresponding column of $C(W)$ contains only zeros, so suppose henceforth that $\Prop(j_m) \neq \emptyset$. In particular, this implies that $m>1$, since if $m=1$ we would have $\Prop(j_1) \subseteq \cP_0 = \emptyset$.

In order to exhibit a dependent subset of vertices in $J$, we first define a sequence of sets of propagators by
\[Q_0 = \Prop(j_m), \qquad Q_{l+1} = \Prop(I_i(Q_l))\quad \forall l \geq 0.\]
It is easily verified that this is an increasing chain of sets: indeed, if $q \in Q_l$ then $I_i(q) \in I_i(Q_l)$, and since $I_i(q)$ is one of the vertices that supports $q$ we must have $q \in \Prop(I_i(Q_l)) = Q_{l+1}$. The sequence therefore stabilizes in finitely many steps, since $\cP$ is finite.

We claim that $Q_l \subseteq \cP_{m-1}$ for all $l\geq 0$. The case $l= 0$ follows immediately from the observation above that $\Prop(j_m) \subseteq \cP_{m-1}$.

Now assume that $Q_0,\dots,Q_{l-1} \subseteq \cP_{m-1}$, and consider $Q_{l}$. If $Q_{l} \setminus Q_{l-1} = \emptyset$ then we are done; if not, let $q \in Q_{l} \setminus Q_{l-1}$. By construction there is some propagator $p_r \in Q_{l-1}$ such that $p_r$ and $q$ are both supported on $I_i(p_r) = a_r$. If $q$ is clockwise of $p_r$ on vertex $a_r$ then we immediately have $q \in \cP_{m-1}$ as observed above. However, if $q$ is counter-clockwise of $p_r$ then we need to show that $p_r$ or another propagator in $Q_l$ constrains the positioning of the other end of $q$ and prevents it from ``escaping'' past $j_m$.

Note that we cannot have $a_r+ 1 = j_m$ when $q$ is counterclockwise of $p_r$, since $q \not\in \Prop(j_m)$. If $V(p_r) \subseteq [I_i(p_r)-1,j_m]$, then since $q$ is inside $p_r$ we also have $V(q) \subseteq [a_r-1,j_m]$ and hence $q \in \cP_{m-1}$ by Lemma [no fourth vertex]. In particular, if $l=1$ then $p_r \in Q_{0} = \Prop(j_m)$ and we must be in this case. If not, then there must be some propagator $p_s \in Q_{l-2}$ such that $p_r$ and $p_s$ are both supported on $I_i(p_s)$; further, we can assume that $q$ is not supported on $I_i(p_s)$ since $q \not\in Q_{l-1}$. This yields one of the configurations in Figure~\ref{fig:awkward induction step}.

\begin{figure}[h]
\includegraphics[scale = 0.3]{awkward_induction_case.jpg}
\caption{The awkward case of the induction step, where $p_r$ does not form a nice arc to constrain $q$. Note that in theory $p_s$ could instead be inside $p_r$ (w.r.t $<_i$), but this does not change the argument and we omit the depiction of this case since it turns out to be impossible.}\label{fig:awkward induction step}
\end{figure}

Now repeat the argument above with $p_s$ in place of $p_r$; this is valid since we only really need to know that $q$ is inside $p_s$ with respect to $<_{a_s - 1}$. Then either $V(p_s) \subseteq [I_i(p_s)-1,j_m]$, in which case we are done (as above), or there exists some $p_t \in Q_{l-3}$ and we continue to iterate the argument. Since there are finitely many propagators and we are guaranteed that $V(p_*) \subseteq [I_i(p_*)-1,j_m]$ if we ever reach $p_* \in Q_0$, we eventually construct a propagator that constrains $q$ and ensures that $q \in \cP_{m-1}$. This completes the induction step.

Finally, let $Q:= \lim_{l\rightarrow \infty} Q_l$; by the above, this is well-defined and contained in $\cP_{m-1}$. We can now show that $J$ is a dependent set, by constructing a subset $U \subseteq J$ that satifies $|\Prop(U)| < |U|$.

Specifically, we take $U = I_i(Q) \cup \{j_m\}$. This is a subset of $J$ because $I_i$ and $J$ agree on their first $m-1$ terms and $Q \subseteq \cP_{m-1}$. Further, we have $j_m \not\in I_i(Q)$, so $|U| = |I_i(Q)| + 1 = |Q| +1$. On the other hand, since $\Prop(j_m) \subseteq Q$, we have
\[\Prop(U) = \Prop(I_i(Q)) \cup \Prop(j_m) = Q\]
by the construction of $Q$, and hence $|\Prop(U)| = |Q| < |U|$. Thus $J$ is a dependent set, and the Grassmann necklace $(I_1,\dots,I_n)$ constructed from $W$ does indeed define the positroid $M(W)$.
\end{proof}

\end{document}

>>>>

\usepackage{amssymb, amsmath, verbatim, amsthm,url, multirow,fullpage,mathtools, outlines}
\usepackage{longtable, rotating,makecell,array}
\usepackage[aligntableaux=top]{ytableau}

@@

Write the matrix $C(W)$ such that the $m^{th}$ row is defined by the propagator $p_m$. We give an algorithm to define a new matrix $D(W)$, where the columns of $D(W)$ are linear combinations of the columns of $C(W)$.

Let $r = 1$
\begin{enumerate}
\item Replace column $\vec{a}_r$ with a normalized version where the entry $(r, a_r) = 1$. 
	\begin{enumerate} \item This can be done since $p_r$ is supported on $a_r$, and subsequent steps to do not set this to $0$. 
	\end{enumerate}
\item For any $e$ greater than $a_r$ ($a_r <_i e$), if $(r, e) = x_{r,e} \neq 0$, replace column $\vec{e}$ with $\vec{e} - x_{r,e} \vec{a}_r$.  
	\begin{enumerate}
	\item Note that this sets every entry of the form $(q, e) = 0$ for $q \leq r$ and $r <_i e$. 
	\item All non-zero entries $(s, e)$, below $(r, e)$ (i.e. $r \leq s$) remain non-zero. In fact, each $x_{s,e}$ is a linear combination of rational functions of the variables $c_{p_r, v}$ that define $C(W)$.
	\end{enumerate}
\item Increment $r$ and repeat.
\end{enumerate}

The algorithm above has a $1$ in every $(r, a_r)$, with $0$s above it and $0$s in all subsequent (in the $<_i$ order) entries in the $r^{th}$ row. It remains to check the columns not in $I_i$.

If $a_k < e$, the $\vec{e}$ column has all $0$ entries by the algorithm. These columns can all be expressed as a linear combination of the columns indicated in $I_i$.

Suppose $a_m + 1 < a_{m+1}$. That is, two consecutive propagators, $p_m$ and $p_{m+1}$ do not contribute consecutive indicies. (Without loss of generality, assume that $m$ is the smallest integer for which this is true.) We claim at the vertices in the cyclic interval $[a_m+1, a_{m+1}-1]$ are also columns of $0$s in $D(W)$. First note that they only support propagators in $\cP_m$. Indeed, if $v \in [a_m+1, a_{m+1}-1]$, and $v$ supported such a propagator, $q \in \Prop(v) \setminus \cP_m$, then $q$ would not have contributed to $I_i$ by the time the algorithm reached $v$, forcing $q$ to contribute either after $p_m$ and before $p_{m+1}$, which is a contradiction, or $q = p_{m+1}$, which contradicts $v < a_{m+1}$.

Therefore, for column $\vec{v}$ to have a non-zero entry, there must exists a $q$ that contributes after $p_m$ supported on an $a_{r_1}$ (with $r_1 < m$) and a sequence $\{a_{r_1} \ldots a_{r_j}, v\}$ such that each contibuting propagator is supported on the next assigned vertex in the series, $p_{r_i} \in \Prop(a_{r_{i+1}})$ and $p_{r_j} \in \Prop(v)$.\sanote{Key condition to check that I haven't missed any cases.} 
We show that this cannot happen. Note that when $p_{r_1}$ contributes $I_i(p_{r_1}) = a_{r_1}$ to the Grassmann necklace, it must be the clockwise-most propagator (amongst those not yet assigned) that is supported on $a_{r_1}$ i.e. clockwise-most amongst propagators in the set $\cP \setminus \cP_{r_1-1}$. Therefore, if $q$ is assigned after $p_{r_1}$ but also supported on $a_{r_1}$, $q$ must be inside $p_{r_1}$ in the $a_{r_1} -1$ ordering. However, we know that $p_{r_1}$ is also supported on $a_{r_2} >_i a_{r_2}$. This either means that $q$ is also supported on $a_{r_2}$, in which case, we may start the above sequence from $a_{r_2}$, or that $q$ contributes a vertex in the interval $[a_{r_1} +1,a_{r_2} -1]$, which implies that $q$ contributes before $p_{m+1}$, which is a contradition.

Therefore, any set of the form $\{a_1 \ldots a_m, v\}$ is linearly dependent in $C(W)$. Thus any $J$ lexicographically smaller than $I_i$ with respect to $<_i$ must contain a set of this form, we show that $J$ is not linearly dependent.
\end{proof}
\end{document}
